---
title: "566hw2"
output: html_document
---
#Download and input the data

```{r}
library(data.table)
library(tidyverse)
library(leaflet)
library(dplyr)
Individual <- data.table::fread("/Users/jiqingwu/Desktop/566 Introduction to Health Data Science/hw2/chs_individual.csv")
Regional <- data.table::fread("/Users/jiqingwu/Desktop/566 Introduction to Health Data Science/hw2/chs_regional.csv")
```


## Part 1
# 1. Merge the datasets

```{r}
chs = merge(
  x = Individual, y = Regional, 
  by.x = "townname", by.y = "townname",
  all.x = TRUE, all.y = TRUE
)

nrow(Individual)
nrow(Regional)
nrow(chs)

```

#There are 1200 rows in the Individual dataset, 12 rows in Regional dataset and 1200 rows in the merged dataset.

# Impute data using the average within the variables “male” and “hispanic.”

```{r}
chs[,mean_bmi:=mean(bmi,na.rm=TRUE),by=.(male, hispanic)]
chs[,bmi:=ifelse(is.na(bmi),mean_bmi,bmi)]

chs[,mean_fev:=mean(fev,na.rm=TRUE),by=.(male, hispanic)]
chs[,fev:=ifelse(is.na(fev),mean_fev,fev)]

chs[,mean_gasstove := round(mean(gasstove,na.rm=TRUE)),by=.(male,hispanic)]
chs[,gasstove:=ifelse(is.na(gasstove),mean_gasstove,gasstove)]

chs[,mean_smoke := round(mean(smoke,na.rm=TRUE)),by=.(male,hispanic)]
chs[,smoke:=ifelse(is.na(smoke),mean_smoke,smoke)]
```

# 2. Create obesity level and a summary table

```{r}
chs$obesity_level <- cut(chs$bmi, 
          breaks=c(-Inf, 14, 22, 24, Inf), 
          labels=c("underweight","normal","overweight","obese"))

chs %>% group_by(obesity_level) %>%
  summarize(Min = min(bmi, na.rm = TRUE), Max = max(bmi, na.rm = TRUE), Count=n()) %>%
  filter(!is.na(obesity_level))
```

# 3. Create smoke gas exposure

```{r}
chs$smoke_log = as.logical(chs$smoke)
chs$gasstove_log = as.logical(chs$gasstove)

chs$smoke_gas_exposure[!chs$smoke_log & !chs$gasstove_log] <-"no smoke or gas exposure"
chs$smoke_gas_exposure[!chs$smoke_log & chs$gasstove_log] <-"gas exposure"
chs$smoke_gas_exposure[chs$smoke_log & !chs$gasstove_log] <-"smoke exposure"
chs$smoke_gas_exposure[chs$smoke_log & chs$gasstove_log] <-"smoke and gas exposure"
```

# 4. Create four summary tables

```{r}
# by town
chs[, .(
  fev_avg = mean(fev, na.rm = TRUE), 
  fev_sd = sd(fev, na.rm = TRUE),
  asthma_prop = paste0(round(100*sum(asthma, na.rm = TRUE)/length(asthma), 2), '%'),
  non_asthma_prop = paste0(round(100-100*sum(asthma, na.rm = TRUE)/length(asthma), 2), '%')
),
by = townname
]

# male
chs[, .(
  fev_avg = mean(fev, na.rm = TRUE), 
  fev_sd = sd(fev, na.rm = TRUE),
  asthma_prop = paste0(round(100*sum(asthma, na.rm = TRUE)/length(asthma), 2), '%'),
  non_asthma_prop = paste0(round(100-100*sum(asthma, na.rm = TRUE)/length(asthma), 2), '%')
),
by = male
]

# obesity level
chs[, .(
  fev_avg = mean(fev, na.rm = TRUE), 
  fev_sd = sd(fev, na.rm = TRUE),
  asthma_prop = paste0(round(100*sum(asthma, na.rm = TRUE)/length(asthma), 2), '%'),
  non_asthma_prop = paste0(round(100-100*sum(asthma, na.rm = TRUE)/length(asthma), 2), '%')
),
by = obesity_level
] %>% filter(!is.na(obesity_level))

# smoke_gas_exposure
chs[, .(
  fev_avg = mean(fev, na.rm = TRUE), 
  fev_sd = sd(fev, na.rm = TRUE),
  asthma_prop = paste0(round(100*sum(asthma, na.rm = TRUE)/length(asthma), 2), '%'),
  non_asthma_prop = paste0(round(100-100*sum(asthma, na.rm = TRUE)/length(asthma), 2), '%')
),
by = smoke_gas_exposure
]%>% filter(!is.na(smoke_gas_exposure))
```

## Part 2
# Check the main variables

```{r}
dim(chs)
head(chs)
tail(chs)

str(chs)

summary(chs$bmi)

summary(chs$fev)

summary(chs$pm25_mass)

table(chs$smoke_gas_exposure)
```

# Check for the dimensions, headers, footers. Missing values are imputed using mean value at the begining.  There is no implausible values in these four main variables. 

# 1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

```{r}
chs %>% 
  filter(!(townname %in% NA)) %>% 
  ggplot(mapping = aes(x = bmi, y = fev)) + 
  geom_point() + 
  geom_smooth(method = 'lm', formula = y~x) +
  facet_wrap(~ townname) +
  ggtitle("Relationship of bmi and fev in different towns")

```

# 2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

```{r}

# by obesity level
chs %>%
  #filter(!(obesity_level %in% NA)) %>%
ggplot(diamonds, mapping = aes(x = fev, fill = obesity_level)) +
  scale_fill_brewer(palette = "GnBu") +
  geom_histogram(binwidth = 20) + 
  labs(title="histograms of FEV by BMI category")

# by smoke/gas exposure 
chs %>%
  #filter(!(smoke_gas_exposure %in% NA)) %>%
ggplot(diamonds, mapping = aes(x = fev, fill = smoke_gas_exposure)) +
  scale_fill_brewer(palette = "GnBu") +
  geom_histogram(binwidth = 20) +
  labs(title="histograms of FEV by smoke/gas exposure")
```

# People with a higher fev seems more likely to have a higher bmi while people with different smoke and gas exposure situation seems to have a more uniform distribution in fev. In other words, bmi has a positive relationship with fev, whille gas/smoke exposure does not affect fev. 

# 3. Barchart of BMI by smoke/gas exposure.

```{r}
chs %>%
  #filter(!(smoke_gas_exposure %in% NA)) %>%
  #filter(!(obesity_level %in% NA)) %>%
  ggplot() +
  geom_bar(mapping = aes(x = obesity_level, fill = smoke_gas_exposure)) +
  scale_fill_brewer(palette = "GnBu") + 
   labs(title="Barchart of BMI level by smoke/gas exposure")
```

# It seems that a larger proportions of obese people are exposured to smoke. Most of people are exposed to gas, and this does not affect obesity level.

# 4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.

```{r}
# by bmi level
chs %>%
  #filter(!(obesity_level %in% NA)) %>% 
  ggplot() + 
  stat_summary(mapping = aes(x = obesity_level, y = fev),
               fun.min = min,
               fun.max = max,
               fun = median) +
labs(title="Summary graphs of fev by obesity level")

# by smoke/gas exposure
chs %>%
  #filter(!(smoke_gas_exposure %in% NA)) %>% 
  ggplot() + 
  stat_summary(mapping = aes(x = smoke_gas_exposure, y = fev),
               fun.min = min,
               fun.max = max,
               fun = median) +
  labs(title="Summary graphs of fev by smoke/gas exposure")
```

# From the summary graphs, people with a higher bmi level are more likely to have a higher fev while people with different gas/smoke exposure situation seem to have a similar fev. 

# 5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r}
pm_pal<-colorNumeric(c("blue","purple","red"), domain=chs$pm25_mass)

leaflet(chs) %>%
  addProviderTiles("OpenStreetMap") %>%
  addCircles(lat=~lat, lng=~lon, color=~pm_pal(pm25_mass),opacity=1,fillOpacity=1,radius=100) %>%
  addLegend("bottomleft", pal=pm_pal, values=chs$pm25_mass, title="PM 2.5 concentration in CHS commuities", opacity=1)

```

# The map shows that red dots (area with higher PM 2.5 concentrations) are concentrated near the city (e.g. LA).

# 6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r}
chs %>% 
  ggplot(mapping = aes(x = pm25_mass, y = fev)) + 
  geom_point() + 
  geom_smooth(method = 'lm', formula = y~x) +
  ggtitle("Relationship of PM 2.5 and fev")

cor.test(chs$pm25_mass, chs$fev)
```

# From the graph above, the regression line shows a slight descending trend, while the change is small. From the correlation test, the p-value=0.01096<0.05. Therefore, the relationship between PM 2.5 and FEV is significant. Correlation coefficient is -0.0734 which indicates a negative relationship. As the concentration of PM 2.5 in an area increase, the FEV value would decrease.

